### 2013-07-23 doc/asplos14/overview.tex
edits mostly to overview

Commodity hardware systems provide fast processors and high-bandwidth
interconnects but suffer when they encounter unpredictable data access
patterns. Maximizing throughput in such a situation depends upon proper
management of concurrency and communication. The \Grappa runtime
components handle thousands of concurrent accesses and batch them to
efficiently use the network. The imbalance between network bandwidth
and core-cycles for communication-intensive codes makes supporting
these functions in software an attractive design point. 
\Grappa specifically employs user-mode threading and networking for
low overhead because context switches and messages are very frequent.

As \Grappa depends on concurrency for performance, its programming
model focuses on enabling programmers to express large amounts of
concurrency. The overall goal is for the system to provide the
critical components for a well-performing implementation of an
irregular application, allowing the programmer to foremost focus on
expressing concurrency. While the programmer can initially be
productive in a shared memory environment and rely on the
underlying runtime components for reasonable performance, \Grappa
provides constructs for optimizing for locality and overlapping
communication.

@@

\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.95\columnwidth]{figs/system-overview}
\begin{minipage}{0.95\columnwidth}
  \caption{\label{fig:grappa} \Grappa system overview}
\end{minipage}
\vspace{-3ex}
\end{center}
\end{figure}

>>>>

% Commodity clusters provide fast processors and high-bandwidth
% interconnects but suffer when they encounter unpredictable data access
% patterns. Maximizing throughput in such a situation depends upon
% proper management of concurrency and communication.

\Grappa's design is motivated by two key observations of today's systems.
First, modern compute cores can execute hundreds of thousands of instructions
in the few microseconds that it takes for one RDMA read to complete. Second,
the injection rate of modern commodity networks is low and the bandwidth is
optimized for large messages. Given the frequent communication caused by high
degree of fine-grained random access in irregular applications, it is
necessary to both overlap communication with computation as much as possible
and to aggregate network messages efficiently. Providing these two properties
depend on exploiting large amounts of concurrency. Hence, \Grappa's
programming model focuses on enabling programmers to easily express
concurrency.

% \Grappa is intended to support what we
% think is a desirable workflow: programmers start by expressing a computation
% using a multithreaded, shared-memory model on a single node. \Grappa then
% provides the primitives to do a straightforward translation of this code to
% run on a distributed-memory cluster, through constructs for optimizing
% locality and overlapping communication.

@@

\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.95\columnwidth]{figs/system-overview}
\begin{minipage}{0.95\columnwidth}
  \caption{\label{fig:grappa} \Grappa system overview\TODO{Cora->Manta}}
\end{minipage}
\vspace{-3ex}
\end{center}
\end{figure}
